---
title: "Trait Plasticity based on climate transfer distance"
author: "Anoob Prakash"
date: "03/03/2021"
output:
  rmarkdown::html_document:
    theme: cosmo  
    number_sections: false
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Methods  

**family BLUP model**  
- Mixed effect model run to get the family level blups for use in the MCMC run.  
- Garden is set as the fixed effect.  
- Family and mBed (5 beds x 3 sites) set as random effect.  
- Height growth for a season is used as the fitness proxy to compare the adaptive nature of trait plasticity.  
  - This helps in explaining whether the change in plasticity for a trait (positive/negative) is adaptive (top right of the plot) or maladaptive (bottom left of the plot).  
  
```
# Creating mBed
fitness <- fitness %>% unite(mBed, Garden, Bed, sep = "_", remove = FALSE)

# lmer model for family blup 
fitness_mod <- lmer(Growth ~ Garden + (1|mBed) + (1|Family), fitness)

# extract the blups for the family from the model  
fitness_blup <- ranef(fitness_mod)
fitness_blup <- fitness_blup$Family
fitness_blup <- cbind(Family=rownames(fitness_blup),fitness_blup)
rownames(fitness_blup) <- NULL
names(fitness_blup)[2] <- "fitness"
```

**Trait plasticity (as described in the paper)**   
To estimate phenotypic plasticity for each trait, we ran LMMs (Model IV, Table S1) that included the fixed effect of garden climate (gPC1, based on each garden’s eigenvector score along PC1 of the climate PCA for the year of measurement), and random effects of Bed (1|Bed) and Family with random intercepts and slopes (gPC1|Family). We used the ‘coef()’ function to extract the random slopes for each Family with gPC1, which provides a measure of plasticity to garden climate (Arnold et al. 2019). To test if the magnitude of plasticity reflected differences due to source climate variability, we modeled the absolute value of plasticity as a function of source climate (sPC1), with a random effect of Population (Model V, Table S1).  
   


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# setwd("Z:/Anoob/MCMCglmm")

suppressPackageStartupMessages({
# packages
require(MCMCglmm)
require(lmerTest)
require(dplyr)
require(tidyr)
require(data.table)
require(ggplot2)
require(cowplot)
require(plyr)
})

# data 
budset_2019 <- read.csv("./trait_data/BudSet_2019.csv")
budset_2020 <- read.csv("./trait_data/BudSet_2020.csv")
budbreak_2020 <- read.csv("./trait_data/BudBreak_2020_cGDD.csv")
heightGrowth_2019 <- read.csv("./trait_data/Growth_2019.csv")
heightGrowth_2020 <- read.csv("./trait_data/Growth_2020.csv")
# Plasticity <- read.csv("./trait_data/Plasticity.csv")
Meta <- read.table("./Exome/RS_Exome_metadata.txt", sep="\t",header=T)

PC_scores <- read.csv("./trait_data/PCA/PCA_score.csv", header=T)
PC_scores <- PC_scores[!is.na(PC_scores$Family),]

# meta data and PC 

colnames(Meta)[colnames(Meta)=="Pop"] <- "Population"
Meta$Region <- as.factor(Meta$Region)
Meta$Region <- mapvalues(Meta$Region, from = c("C", "M", "E"), to = c("Core", "Margin", "Edge"));Meta$Region <- factor(Meta$Region,levels = c("Core", "Margin", "Edge"))


Meta <- merge(x=Meta,
                     y=PC_scores[,-c(1,2,3,5,17)],
                     by.x="Family",
                     by.y="Family")

Meta <- Meta[,c("Population","Family","Region","Latitude","Longitude","Elevation","PC1","PC2","PC3","PC4","PC5","PC6",
                "PC7","PC8","PC9","PC10","PC11")]
```

# Garden climate  

**Garden climate**  
```{r}
Garden_clim <- read.table("./ClimateNA/climNA_gardens_2019-20.txt", header=T)
Garden_clim
selVars1 <- c('DD_0','DD18','MAR','PAS','MSP','RH','EXT','CMD','TD','eFFP','PET')
colnames(Garden_clim)[1] <- "Garden"
colnames(Garden_clim)[2] <- "Year"
```

```{r}
# create ID for garden means
# garden_identity <- Garden_climateNA1[,1]

# create ID for garden
garden_identity <- Garden_clim[,1:2]

# selected clim of gardens
data_p1 <- Garden_clim[,selVars1]
```

```{r}
garden_id <- garden_identity

colnames(garden_id)[1] <- "Garden"
colnames(garden_id)[2] <- "Year"

data_pca <- prcomp(Garden_clim[,-c(1:5)],scale. = T)

# Extract PC axes for plotting
PCAvalues <- data.frame(Garden = Garden_clim$Garden, Year = Garden_clim$Year,
                        data_pca$x)
PCAvalues$Year <- as.factor(PCAvalues$Year)

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(data_pca$rotation), data_pca$rotation)

# write.csv(PCAvalues, row.names = F, "./trait_data/PCA/PCA_garden_score.csv")
# write.csv(PCAloadings, row.names = F, "./trait_data/PCA/PCA_garden_loadings.csv")

PCAvalues_mod <- PCAvalues
PCAvalues_mod$Garden[PCAvalues_mod$Garden=="VT"] <- "Vermont"
PCAvalues_mod$Garden[PCAvalues_mod$Garden=="MD"] <- "Maryland"
PCAvalues_mod$Garden[PCAvalues_mod$Garden=="NC"] <- "North_Carolina"

```




# Trait plasticity   {.tabset .tabset-fade .tabset-pills}  

## Bud break 2020  

- cGDD  

```{r}
budbreak_2020 <- merge(x=budbreak_2020,y=PCAvalues_mod[PCAvalues_mod$Year=="2020",], all.x=TRUE)
budbreak_2020$mBed <- paste0(budbreak_2020$Garden,"_",budbreak_2020$Bed)

#plasticity model
budbreak_PC1_mod <- lmer(data = budbreak_2020, na.action = na.omit,
                          cGDD ~ PC1  + (1|mBed) + (PC1|Family))

summary(budbreak_PC1_mod)

budbreak_PC1 <- ranef(budbreak_PC1_mod)
budbreak_PC1 <- budbreak_PC1$Family
budbreak_PC1 <- cbind(Family=rownames(budbreak_PC1),budbreak_PC1)
rownames(budbreak_PC1) <- NULL

budbreak_PC1$Family <- as.factor(budbreak_PC1$Family)
colnames(budbreak_PC1)[colnames(budbreak_PC1)=="PC1"] <- "BudBreak_2020"
budbreak_PC1 <- budbreak_PC1[,-2]
head(budbreak_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

bb2020x_PC1 <- coef(budbreak_PC1_mod)
# hist(testx_PC1$PC1)
bb2020x_PC1 <- bb2020x_PC1$Family
bb2020x_PC1 <- cbind(Family=rownames(bb2020x_PC1),bb2020x_PC1)
rownames(bb2020x_PC1) <- NULL
bb2020x_PC1$abs <- abs(bb2020x_PC1$PC1)
hist(bb2020x_PC1$PC1)
hist(bb2020x_PC1$abs)
bb2020x_PC1 <- bb2020x_PC1[,-2]
colnames(bb2020x_PC1)[3] <- "BudBreak_cGDD"
bb2020x_PC1 <- bb2020x_PC1[,-2]  
```

- DOY  

```{r}
budbreak_2020_DOY <- merge(x=budbreak_2020,y=PCAvalues_mod[PCAvalues_mod$Year=="2020",], all.x=TRUE)
budbreak_2020_DOY$mBed <- paste0(budbreak_2020_DOY$Garden,"_",budbreak_2020_DOY$Bed)

#plasticity model
budbreak_DOY_PC1_mod <- lmer(data = budbreak_2020_DOY, na.action = na.omit,
                          BudBreak ~ PC1  + (1|mBed) + (PC1|Family))

summary(budbreak_DOY_PC1_mod)

budbreak_DOY_PC1 <- ranef(budbreak_DOY_PC1_mod)
budbreak_DOY_PC1 <- budbreak_DOY_PC1$Family
budbreak_DOY_PC1 <- cbind(Family=rownames(budbreak_DOY_PC1),budbreak_DOY_PC1)
rownames(budbreak_DOY_PC1) <- NULL

budbreak_DOY_PC1$Family <- as.factor(budbreak_DOY_PC1$Family)
colnames(budbreak_DOY_PC1)[colnames(budbreak_DOY_PC1)=="PC1"] <- "BudBreak_2020_DOY"
budbreak_DOY_PC1 <- budbreak_DOY_PC1[,-2]
head(budbreak_DOY_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

bb2020_DOYx_PC1 <- coef(budbreak_DOY_PC1_mod)
# hist(testx_PC1$PC1)
bb2020_DOYx_PC1 <- bb2020_DOYx_PC1$Family
bb2020_DOYx_PC1 <- cbind(Family=rownames(bb2020_DOYx_PC1),bb2020_DOYx_PC1)
rownames(bb2020_DOYx_PC1) <- NULL
bb2020_DOYx_PC1$abs <- abs(bb2020_DOYx_PC1$PC1)
hist(bb2020_DOYx_PC1$PC1)
hist(bb2020_DOYx_PC1$abs)
bb2020_DOYx_PC1 <- bb2020_DOYx_PC1[,-2]
colnames(bb2020_DOYx_PC1)[3] <- "BudBreak_DOY"
bb2020_DOYx_PC1 <- bb2020_DOYx_PC1[,-2]  
```

## Bud set 2019 

```{r}
budset_2019 <- merge(x=budset_2019,y=PCAvalues_mod[PCAvalues_mod$Year=="2019",], all.x=TRUE)
budset_2019$mBed <- paste0(budset_2019$Garden,"_",budset_2019$Bed)

# plasticity model
budset_2019_PC1_mod <- lmer(data = budset_2019, na.action = na.omit,
                          BudSet ~ PC1  + (1|mBed) + (PC1|Family))

summary(budset_2019_PC1_mod)

budset_2019_PC1 <- ranef(budset_2019_PC1_mod)
budset_2019_PC1 <- budset_2019_PC1$Family
budset_2019_PC1 <- cbind(Family=rownames(budset_2019_PC1),budset_2019_PC1)
rownames(budset_2019_PC1) <- NULL

budset_2019_PC1$Family <- as.factor(budset_2019_PC1$Family)
colnames(budset_2019_PC1)[colnames(budset_2019_PC1)=="PC1"] <- "BudSet_2019"
budset_2019_PC1 <- budset_2019_PC1[,-2]

bs2019x_PC1 <- coef(budset_2019_PC1_mod)
# hist(testx_PC1$PC1)
bs2019x_PC1 <- bs2019x_PC1$Family
bs2019x_PC1 <- cbind(Family=rownames(bs2019x_PC1),bs2019x_PC1)
rownames(bs2019x_PC1) <- NULL
bs2019x_PC1$abs <- abs(bs2019x_PC1$PC1)
hist(bs2019x_PC1$PC1)
hist(bs2019x_PC1$abs)
bs2019x_PC1 <- bs2019x_PC1[,-c(2:3)]
colnames(bs2019x_PC1)[2] <- "BudSet_2019"  
```

## Bud set 2020  

```{r}
budset_2020 <- merge(x=budset_2020,y=PCAvalues_mod[PCAvalues_mod$Year=="2020",], all.x=TRUE)
budset_2020$mBed <- paste0(budset_2020$Garden,"_",budset_2020$Bed)



# plasticity model
budset_2020_PC1_mod <- lmer(data = budset_2020, na.action = na.omit,
                          BudSet ~ PC1  + (1|mBed) + (PC1|Family))

summary(budset_2020_PC1_mod)

budset_2020_PC1 <- ranef(budset_2020_PC1_mod)
budset_2020_PC1 <- budset_2020_PC1$Family
budset_2020_PC1 <- cbind(Family=rownames(budset_2020_PC1),budset_2020_PC1)
rownames(budset_2020_PC1) <- NULL

budset_2020_PC1$Family <- as.factor(budset_2020_PC1$Family)
colnames(budset_2020_PC1)[colnames(budset_2020_PC1)=="PC1"] <- "BudSet_2020"
budset_2020_PC1 <- budset_2020_PC1[,-2]

bs2020x_PC1 <- coef(budset_2020_PC1_mod)
# hist(testx_PC1$PC1)
bs2020x_PC1 <- bs2020x_PC1$Family
bs2020x_PC1 <- cbind(Family=rownames(bs2020x_PC1),bs2020x_PC1)
rownames(bs2020x_PC1) <- NULL
bs2020x_PC1$abs <- abs(bs2020x_PC1$PC1)
hist(bs2020x_PC1$PC1)
hist(bs2020x_PC1$abs)

bs2020x_PC1 <- bs2020x_PC1[,-c(2:3)]
colnames(bs2020x_PC1)[2] <- "BudSet_2020"
```

## Growth 2019 


```{r}
growth_2019 <- merge(x=heightGrowth_2019,y=PCAvalues_mod[PCAvalues_mod$Year=="2019",], all.x=TRUE)
growth_2019$mBed <- paste0(growth_2019$Garden,"_",growth_2019$Bed)



# plasticity model
growth_2019_PC1_mod <- lmer(data = growth_2019, na.action = na.omit,
                          Growth ~ PC1  + (1|mBed) + (PC1|Family))

summary(growth_2019_PC1_mod)

# ranef
growth_2019_PC1 <- ranef(growth_2019_PC1_mod)
growth_2019_PC1 <- growth_2019_PC1$Family
growth_2019_PC1 <- cbind(Family=rownames(growth_2019_PC1),growth_2019_PC1)
rownames(growth_2019_PC1) <- NULL

growth_2019_PC1$Family <- as.factor(growth_2019_PC1$Family)
colnames(growth_2019_PC1)[colnames(growth_2019_PC1)=="PC1"] <- "Growth_2019"
growth_2019_PC1 <- growth_2019_PC1[,-2]

# coef
gw2019x_PC1 <- coef(growth_2019_PC1_mod)
gw2019x_PC1 <- gw2019x_PC1$Family
gw2019x_PC1 <- cbind(Family=rownames(gw2019x_PC1),gw2019x_PC1)
rownames(gw2019x_PC1) <- NULL
gw2019x_PC1$abs <- abs(gw2019x_PC1$PC1)
hist(gw2019x_PC1$PC1)
hist(gw2019x_PC1$abs)

gw2019x_PC1 <- gw2019x_PC1[,-c(2:3)]
colnames(gw2019x_PC1)[2] <- "Growth_2019"  
```

## Growth 2020  

```{r}
growth_2020 <- merge(x=heightGrowth_2020,y=PCAvalues_mod[PCAvalues_mod$Year=="2020",], all.x=TRUE)
growth_2020$mBed <- paste0(growth_2020$Garden,"_",growth_2020$Bed)



# plasticity model
growth_2020_PC1_mod <- lmer(data = growth_2020, na.action = na.omit,
                          Growth ~ PC1  + (1|mBed) + (PC1|Family))

summary(growth_2020_PC1_mod)

growth_2020_PC1 <- ranef(growth_2020_PC1_mod)
growth_2020_PC1 <- growth_2020_PC1$Family
growth_2020_PC1 <- cbind(Family=rownames(growth_2020_PC1),growth_2020_PC1)
rownames(growth_2020_PC1) <- NULL

growth_2020_PC1$Family <- as.factor(growth_2020_PC1$Family)
colnames(growth_2020_PC1)[colnames(growth_2020_PC1)=="PC1"] <- "Growth_2020"
growth_2020_PC1 <- growth_2020_PC1[,-2]

gw2020x_PC1 <- coef(growth_2020_PC1_mod)
# hist(testx_PC1$PC1)
gw2020x_PC1 <- gw2020x_PC1$Family
gw2020x_PC1 <- cbind(Family=rownames(gw2020x_PC1),gw2020x_PC1)
rownames(gw2020x_PC1) <- NULL
gw2020x_PC1$abs <- abs(gw2020x_PC1$PC1)
hist(gw2020x_PC1$PC1)
hist(gw2020x_PC1$abs)

gw2020x_PC1 <- gw2020x_PC1[,-c(2:3)]
colnames(gw2020x_PC1)[2] <- "Growth_2020"
```

# Plasticity  
```{r}

Plasticity <- Reduce(function(x, y) merge(x, y, all=TRUE),
                     list(Meta,bb2020x_PC1,bb2020_DOYx_PC1,bs2019x_PC1,bs2020x_PC1,gw2019x_PC1,gw2020x_PC1))
Plasticity <- Plasticity[!is.na(Plasticity$BudBreak_cGDD),]

# write.csv(Plasticity, "./trait_data/Plasticity.csv")
```




